---
title: "Returns on Financial Stocks Analysis"
author: "MAM Group 5"
date: "`r Sys.Date()`"
tags: ["data_visualisation"]
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

A group project calculating returns and examining the distribution of returns based on a stock price dataset

<!--more-->


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest) # to scrape wikipedia page
```

# Returns of financial stocks

We will use the `tidyquant` package to download historical data of stock prices, calculate returns, and examine the distribution of returns.

We must first identify which stocks we want to download data for, and for this we must know their ticker symbol; Apple is known as AAPL, Microsoft as MSFT, McDonald's as MCD, etc. The file `nyse.csv` contains 508 stocks listed on the NYSE, their ticker `symbol`, `name`, the IPO (Initial Public Offering) year, and the sector and industry the company is in.

```{r load_nyse_data, message=FALSE, warning=FALSE}
nyse <- read_csv(here::here("data","nyse.csv"))

glimpse(nyse)

```

Based on this dataset, we have created a table and a bar plot that shows the number of companies per sector, in descending order

```{r companies_per_sector}

# a easier way
# nyse %>% 
#   select(sector) %>% 
#   table() %>%
#   sort(decreasing = T) %>%
#   barplot()

nyse %>%
  group_by(sector) %>% 
  mutate(company_num = count(sector)) %>%
  ggplot(aes(x=company_num, y=fct_reorder(sector, company_num))) +
  geom_bar(stat="identity") +
  labs(title = "Number of Companies in each Sector",
       x = "",
       y = "Sector")

```

Next, we have chosen some stocks and their ticker symbols and downloaded data.
The stocks we chose are: "EBR","BBL","AEE","BCE","BRO","CAT","BUD","SPY"

```{r get_price_data, message=FALSE, warning=FALSE, cache=TRUE}
# Notice the cache=TRUE argument inthe chunk options. Because getting data is time consuming, 
# cache=TRUE means that once it downloads data, the chunk will not run again next time you knit your Rmd

myStocks <- c("EBR","BBL","AEE","BCE","BRO","CAT","BUD","SPY" ) %>%
  tq_get(get  = "stock.prices",
         from = "2011-01-01",
         to   = "2021-08-31") %>%
  group_by(symbol) 

glimpse(myStocks) # examine the structure of the resulting data frame
```

We can see that the dataset is 8x21,464 tibble with each row being the ohlc (open,high,low,close) and volume for a stock on a given date.

Financial performance analysis depend on returns; If I buy a stock today for 100 and I sell it tomorrow for 101.75, my one-day return, assuming no transaction costs, is 1.75%. So given the adjusted closing prices, our first step is to calculate daily and monthly returns.

```{r calculate_returns, message=FALSE, warning=FALSE, cache=TRUE}
#calculate daily returns
myStocks_returns_daily <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "daily", 
               type       = "log",
               col_rename = "daily_returns",
               cols = c(nested.col))  

#calculate monthly  returns
myStocks_returns_monthly <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "monthly_returns",
               cols = c(nested.col)) 

#calculate yearly returns
myStocks_returns_annual <- myStocks %>%
  group_by(symbol) %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "yearly", 
               type       = "arithmetic",
               col_rename = "yearly_returns",
               cols = c(nested.col))
```

We have created a table summarising monthly returns for each of the stocks and `SPY`; min, max, median, mean, SD.

```{r summarise_monthly_returns}
glimpse(myStocks_returns_monthly)

monthlystocks_summarised <- myStocks_returns_monthly %>% 
  group_by(symbol) %>%
  summarise(min_return = min(monthly_returns),
            max_return = max(monthly_returns),
            median_return = median(monthly_returns),
            mean_return = mean(monthly_returns),
            sd_return = sd(monthly_returns))
```

Plotted a density plot, using `geom_density()`, for each of the stocks

```{r density_monthly_returns}

ggplot(myStocks_returns_monthly, aes(x = monthly_returns)) +
  geom_density() + 
  facet_wrap(~symbol) +
  labs(
    title = "Density Plots for Monthly Returns of Stocks",
    x = "Monthly Return",
    y = "Density"
  ) +
  NULL

```


> Inferences from the plots

The monthly returns for the flatter density plots are more dispersed whereas those with tall peaks are more concentrated. The riskiest stock is EBR and the least risky is SPY (as an ETF) due to the shape of their peaks.

Finally, make a plot that shows the expected monthly return (mean) of a stock on the Y axis and the risk (standard deviation) in the X-axis. Please use `ggrepel::geom_text_repel()` to label each stock

```{r risk_return_plot}

monthlystocks_summarised%>%
  ggplot(aes(y = mean_return, x=sd_return)) +
  geom_point() + 
  ggrepel::geom_text_repel(aes(label = symbol)) +
  NULL
   

```


> Inferences from the plots

EBR is the most risky as it has the highest standard deviation of returns. EBR has the highest sd but also a high expected return. AEE, CAT and SPY also have high expected returns with less standard deviation. BBL and BUD have lower expected returns and higher standard deviations meaning they are riskier and do not have high expected returns. SPY produces a good return with lower risk.


# Team members

- Alex Kubbinga
- Clara Moreno Sanchez
- Jean Huang
- Raghav Mehta
- Raina Doshi
- Yuan Gao