---
title: "Alcoho1 Consumption Analysis"
author: "LBS-MAM22 Group 5"
date: "`r Sys.Date()`"
tags: ["groupwork", "data_visualisation"]
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest) # to scrape wikipedia page
```

# Where Do People Drink The Most Beer, Wine And Spirits?

Back in 2014, [fivethiryeight.com](https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/) published an article on alchohol consumption in different countries. The data `drinks` is available as part of the `fivethirtyeight` package. 

```{r, load_alcohol_data}
library(fivethirtyeight)
data(drinks)

```

After skimming the data, we can see that there are no missing values in the dataset.
The variable types are:

+---------------+-----------+---------------+-----------------+---------------+------------------------------+
| Column name   | country   | beer_servings | spirit_servings | wine_servings | total_litres_of_pure_alcohol |
+===============+===========+===============+=================+===============+==============================+
| Variable type | character | integer       | integer         | integer       | double                       |
+---------------+-----------+---------------+-----------------+---------------+------------------------------+

```{r glimpse_skim_data}

glimpse(drinks)
skim(drinks)
```

## Top 25 alcohol consuming countries

We will make a plot to show the top 25 alcohol consuming countries for beer, wine and spirit respectively.

### Top 25 beer consuming countries

```{r beer_plot}
drinks %>% 
  slice_max ( order_by = beer_servings, n=25 ) %>% # taking top 25 countries by servings
  #plotting a graph 
  ggplot(aes(x = beer_servings, y = fct_reorder(country, beer_servings))) +
  geom_col(fill="orange") +
  #labelling the graph and the axes
  labs(
    title = "Top 25 Beer Consuming Countries in 2010",
    subtitle = "Standard Servings Per Person",
    x = "Beer Servings (in cans)",
    y = "Country"
  )
```
### Top 25 wine consuming countries

```{r wine_plot}
drinks %>% 
  slice_max ( order_by = wine_servings, n=25 ) %>% # taking top 25 countries by servings
  ggplot(aes(x = wine_servings, y = fct_reorder(country, wine_servings))) +
  geom_col(fill="dark red") +
  labs(
    title = "Top 25 Wine Consuming Countries in 2010",
    subtitle = "Standard Servings Per Person",
    x = "Wine Servings (in glasses)",
    y = "Country"
  )
```

### Top 25 spirit consuming countries

```{r spirit_plot}
drinks %>% 
  slice_max ( order_by = spirit_servings, n=25 ) %>% # taking top 25 countries by servings
  ggplot(aes(x = spirit_servings, y = fct_reorder(country, spirit_servings))) +
  geom_col(fill="grey") +
  labs(
    title = "Top 25 Spirit Consuming Countries in 2010",
    subtitle = "Servings (in shots) Per Person",
    x = "Spirit Servings",
    y = "Country"
  )

```
> Inferences & Key takeaways-

1.  European countries are high consumers of wine.

2.  Beer is more evenly distributed around the world in the top 25 countries, as compared to wine and spirit.

3.  European countries are higher ranked for overall consumption of beer, wine and spirit.

4. We see the countries with higher population are lower on the graphs, since the data is plotted per population. Hence the distribution of ages in the population within each country will also affect the alcohol, wine and spirits consumption. 

# Team members

- Alex Kubbinga
- Clara Moreno Sanchez
- Jean Huang
- Raghav Mehta
- Raina Doshi
- Yuan Gao
