---
title: "Gapminder Data Analysis"
author: "LBS-MAM22 Group 5"
date: "`r Sys.Date()`"
tags: ["groupwork","data_visualisation","statistics"]
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(infer)
```

# Gapminder revisited

Now we look at multiple data frames listed below-

1. Life expectancy at birth (life_expectancy_years.csv)
2. GDP per capita in constant 2010 US$ (https://data.worldbank.org/indicator/NY.GDP.PCAP.KD)
3.Female fertility: The number of babies per woman (https://data.worldbank.org/indicator/SP.DYN.TFRT.IN)
4. Primary school enrollment as % of children attending primary school (https://data.worldbank.org/indicator/SE.PRM.NENR)
5. Mortality rate, for under 5, per 1000 live births (https://data.worldbank.org/indicator/SH.DYN.MORT)
6. HIV prevalence (adults_with_hiv_percent_age_15_49.csv): The estimated number of people living with HIV per 100 population of age group 15-49.


```{r, get_data, cache=TRUE}

# load gapminder HIV data
hiv <- read_csv(here::here("data","adults_with_hiv_percent_age_15_49.csv"))
life_expectancy <- read_csv(here::here("data","life_expectancy_years.csv"))

# get World bank data using wbstats
indicators <- c("SP.DYN.TFRT.IN","SE.PRM.NENR", "SH.DYN.MORT", "NY.GDP.PCAP.KD")


library(wbstats)

worldbank_data <- wb_data(country="countries_only", #countries only- no aggregates like Latin America, Europe, etc.
                          indicator = indicators, 
                          start_date = 1960, 
                          end_date = 2016)

# get a dataframe of information regarding countries, indicators, sources, regions, indicator topics, lending types, income levels,  from the World Bank API 
countries <-  wbstats::wb_cachelist$countries

```

First, we need to join the three datasets to a complete csv file.

```{r, HIV Table}

hiv_long <- hiv %>% 
  pivot_longer(cols = 2:34, #columns 2 to 34
               names_to = "Year",
               values_to = "HIV_Value")

skim(hiv_long)
hiv_long


life_expectancy_long <- life_expectancy %>% 
                          pivot_longer(cols = 2:302, #columns 2 to 302
                       names_to = "Year",
                    values_to = "Life_Expectancy_Value")

skim(life_expectancy_long)

life_expectancy_long

#Merged HIV & Life_Expectancy data by matching Year and Country
hiv_life_expectancy <- hiv_long %>% inner_join(  life_expectancy_long , by = c ( "country","Year" ))

#Renaming date column to Year on World Bank Data

worldbank_data <- rename(worldbank_data ,Year=date)


worldbank_data <- worldbank_data %>% 
  mutate(Year = as.character(Year))


#skim(hiv_life_expectancy)
#skim(worldbank_data)

#Merging Life Exp & HIV data with World bank data by Country & Year
hiv_life_expectancy_worldbank_data <- hiv_life_expectancy %>% inner_join( worldbank_data , by = c ( "country","Year" )  )

```

We now check the relationship between HIV prevalence and life expectancy by generating a scatterplot with a smoothing line to report the results.

``` {r, scatterplot}




hiv_life_expectancy_worldbank_data <- hiv_life_expectancy_worldbank_data %>% 
  #create new variable 'interval', and assign values based on criteria below:
  mutate(interval = case_when(
    Year %in% c(1980:1989) ~ "1979-1989",
    Year %in% c(1990:2000) ~ "1990-2000",
    Year %in% c(2001:2010) ~ "2001-2010",
    #Year %in% c(1999:2009) ~ "1999-2009",
    TRUE ~ "2011-present"
  ))



ggplot(hiv_life_expectancy_worldbank_data, aes(x = HIV_Value , y = Life_Expectancy_Value)) +
  geom_point() +
  geom_smooth(method = "lm") +
   labs(
    title = "Relationship of HIV vs Life Expectancy",
    x = "HIV Value",
    y = "Life Expectancy"
  )+
  facet_wrap(~interval)+
  NULL


```


We have faceted the graphs by intervals. We see that as the HIV Value increases the life expectancy decreases. 

Now we analyse the relationship between fertility rate and GDP per capita by generating a scatterplot with a smoothing line.

``` {r}


library(countrycode)

#Creating a dataset hiv_life_expectancy_worldbank_data_continent which has a column for continents corresponding to each country

hiv_life_expectancy_worldbank_data_continent <- hiv_life_expectancy_worldbank_data

hiv_life_expectancy_worldbank_data_continent <- cbind(hiv_life_expectancy_worldbank_data, new_col = "continent") 


hiv_life_expectancy_worldbank_data_continent$continent <- countrycode(sourcevar = hiv_life_expectancy_worldbank_data_continent$country,
                            origin = "country.name",
                            destination = "continent")


ggplot(hiv_life_expectancy_worldbank_data_continent, aes(x = NY.GDP.PCAP.KD , y = SP.DYN.TFRT.IN)) +
  geom_point() +
  geom_smooth(method = "lm") +
   labs(
    title = "Relationship",
    x = "GDP",
    y = "Fertility"
  )+
  facet_wrap(~continent)+

  NULL




```

- As we can see from plotting the gdppercapita against the fertility rate, when GDP per capita increases the fertility rate decreases substantially and reaches an upper bound at a fertility rate of 3

- In Asia, Africa, Americas and Ocenia we see the trend that the higher the GDP the lower is the Fertility rate

- However, in Europe we see that the higher GDP countries also have a higher fertility rate


Now we check which regions have the most observations with missing HIV data and generate a bar chart, in descending order.

```{r}


#aggregate(x = HIV_Value, data=hiv_life_expectancy_worldbank_data, count(is.na(x)))


hiv_missing <- hiv_life_expectancy_worldbank_data %>%
  mutate(na_yes_no = ifelse( is.na(HIV_Value) , "Yes" , "No" )  ) %>% 
  group_by(country,na_yes_no) %>%
  summarise(count_missing=n()) %>%
  filter( na_yes_no == "Yes" ) %>%
  arrange(-count_missing)


hiv_missing[1:15,] %>%  
#  slice_max ( order_by = count_missing, n=5 ) %>%
  ggplot(aes(x = count_missing, y = fct_reorder(country, count_missing))) +
  geom_col(fill="orange")+
  labs(x="No. of missing values" , y="Country" , title= "Missing values of HIV data per country")+
    NULL


```


4. Now we see how the mortality rate for under 5 has changed by region. 

In each region, find the top 5 countries that have seen the greatest improvement, as well as those 5 countries where mortality rates have had the least improvement or even deterioration.

``` {r}

#now we filter the data from year 2011 and 1979
mortality_1979_2011 <- hiv_life_expectancy_worldbank_data_continent %>% 
                      filter ( as.numeric(Year) %in% c(  1979,2011 ) )

#now we select the columns to consider
mortality_1979_2011 <- select( mortality_1979_2011 , c( "Year", "continent" , "country" , "SH.DYN.MORT" ))

#we rename the columns to add a character 
mortality_1979_2011$Year = paste("y",mortality_1979_2011$Year,sep="")

#now we make the table wider to allow us to select the difference in values later
mortality_1979_2011_wide <- pivot_wider( data=mortality_1979_2011 , names_from = Year, values_from = SH.DYN.MORT  )


#now we add an additional column to calcualte the difference in mortality values
mortality_1979_2011_wide <- mortality_1979_2011_wide %>% mutate(mortality_diff = y2011 - y1979 ) 

#we group by the region and rank the mortality differences
mortality_1979_2011_ranked <- mortality_1979_2011_wide %>% 
                          group_by(continent) %>% 
                          summarise( country=country,  asc_ranking = rank(mortality_diff) , dsc_ranking = rank(-mortality_diff)  )


#as per our analysis the lesser mortality range difference means the worst improvement and the largest mortality rate differnce in 1979 and 2011 shows the best improved country

mortality_best5 <- mortality_1979_2011_ranked %>% filter( asc_ranking <=5  )

mortality_worst5 <- mortality_1979_2011_ranked %>% filter( dsc_ranking <=5  )


mortality_worst5 <-select( mortality_worst5  , c ( "continent" , "country"))

mortality_worst5


mortality_best5 <-select( mortality_worst5  , c ( "continent" , "country"))

mortality_best5


```

5. We now check the relationship between primary school enrollment and fertility rate.

```{r}

ggplot(hiv_life_expectancy_worldbank_data_continent, aes(x = SP.DYN.TFRT.IN , y = SE.PRM.NENR)) +
  geom_point() +
  geom_smooth(method = "lm") +
   labs(
    title = "Relationship",
    x = "Fertility",
    y = "School Enrollment"
  )+
  facet_wrap(~continent)+
  NULL




```

- Once again we see that Europe and the rest of the continents have different trends.
In Asia, Afirca, Americas and Oceania - the higher the fertility, the lesser is the school enrollment. 

- In Europe, the higher the fertility rate the more is the school enrollment. 

# Team members

- Alex Kubbinga
- Clara Moreno Sanchez
- Jean Huang
- Raghav Mehta
- Raina Doshi
- Yuan Gao
